### 2014-09-03 constraints.tex
changed structure
tech

\section{Exploration of Unexecuted Branches}

We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.

\paragraph{Relaxation of Branching Decisions}

Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace. This is achieved via a symbolic representation of the input trace. Given branching event $e_b$,
\begin{enumerate}
	\item model the branching condition symbolically to execute along the negation of the original branch;
	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
	\item for each suffix $t^s$,
	\begin{enumerate}
		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
	\end{enumerate}
\end{enumerate}
Constraint solving is applied to each of the resulting traces analogously to the original trace.

We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation. \tool\ currently fails if (i) the unexplored branch containts loops or (ii) there are object references that cannot be fully resolved at the branching point. In case of failure, \tool\ moves on to other branches. We demonstrate in Section \ref{sec:eval} that despite these limitations, the increase in coverage thanks to exploration of new paths is significant.

\paragraph{Proof of Correctness}

\input{proof}

>>>>

\section{Exploration of Unexecuted Branches}
We now switch to the second feature of \tool, which is its ability to reason about unexplored branches.
At the high level, we simply need to derive a trace that contains the unexplored branch, after that, we can reduce the analysis to the analysis in Section~\ref{sec:relax} (starting from the SSA processing).  \tool\ derives such traces by symbolically executing the unexplored branches to record the events and including these events in the new traces.

%TODO | and projection, abuse the terms?
As shown in the following, suppose the symbolic execution  $symEngine(\tau, e)$ returns the single-thread traces starting from the branch $e$, which now takes a different decision. The resultant trace is computed by first removing the events thread-locally after $e$ (inclusively) and then appending one of the single-thread traces that represents a path in the unexplored branch of $e$. The negation $neg(e)$ of $e$, which is the same as $e$ except the evaluation result of the boolean expression differs, is also appended.

\begin{algorithmic}[3]
\For {$e: \tau$}
  \If {$type(e)=branch$}
    \State $S=symEngine(\tau, e)$
	\For {$\tau': S$}
	\State $\tau_r=\tau-((\tau|t^e)|>=e)$
	\State $\tau_r=\tau_r.replace(e, neg(e) + \tau')$
	\EndFor
 \EndIf
\EndFor
\end{algorithmic}

{\bf Symbolic Execution\ } Our symbolic execution is realized on top of the dataflow analysis, which generates events for each instruction within the unexplored branch.
The symbolic execution has known limitations in reasoning about the schedules and loops (and recursion). Therefore, our analysis assumes all the events in the unexplored branch happen atomically at the branch event $e$ without interleavings from other threads, as illustrated also in the above algorithm (line 6). In this way, our analysis adopts the sequential reasoning. \tool\ will reschedule the events based on the relaxation. As for loop (or recursion), the analysis terminates the current data flow immediately after one iteration, i.e., we expand the loop for only once. Otherwise, the analysis terminates the data flow normally if the flow goes out of the scope of the unexplored branch.

The analysis starts with the event $e$ (exclusively), with the current state $\sigma$ at $e$ as the initial state. 
Remember that we use the runtime values for two types of events: the heap value of the base objects and the evaluation result of the boolean expression in the branch. 
Other values are maintained but not used. Therefore, we also need to maintain the heap value and the boolean expression value in the symbolic execution.

First, we propose standard kill/gen rules for three basic heap instructions to maintain the heap update along a given path. Specially, we separate the object creation from the local assignment. In the object creation, we assign a unique integer to represent the newly created object and store it into the state $\sigma$. For field reference, $\sigma[x.f]$ is actually short for two-level interpretation $\sigma[\sigma[x].f]$, where the first finds out the base object $\sigma[x]$ and the second searches for the object referenced by its field $f$. The kill/gen rules take effect only when $x.f$ is not of the primitive type.

The analysis also generates the events, where the heap values are stored in the map of the event for latter use. For the branch, we store its boolean evaluation result.
As the boolean result is available only after the branch takes the decision, we delay the generation of the branch event until the first event after the branch.

\begin{table*}
\centering
%\begin{small}
\begin{tabular}{l|c|c}
\multicolumn{1}{c|}{Operation} & {Kill} & {Gen} \\
\hline
{\tt $y$=$new (...)$} 			&  $\sigma[y] \mapsto \star$		&  $\sigma[y] \mapsto newI$, $<t, cnt, y=newI, \bot>$\\
{\tt $y$=$x.f$} 			&   $\sigma[y]\mapsto \star$			&  $\sigma[y] \mapsto \sigma[x.f]$, $<t, cnt, y=x.f, \lsyn x \rsyn=\sigma[x]>$  \\
{\tt $x.f=y$} & $\sigma[x.f] \mapsto \star$		&  $\sigma[x.f] \mapsto \sigma[y]$, $<t, cnt, x.f=y, \lsyn x \rsyn=\sigma[x]>$ \\
{\tt $z=x\  bop\  y$} & 		&  $<t, cnt, z=x\  bop\  y, \bot>$ \\
{\tt $if(x<y)$} & 	  &  $<t, cnt, if(x<y), \lsyn x<y \rsyn=bool>$\\
\end{tabular}
\caption{\label{table:killgen}Dataflow analysis}
%\end{small}
\end{table*}

%Beyond relaxing flow dependencies, \tool\ is also capable of exploring code branches that were not executed in the original trace. This is achieved via a symbolic representation of the input trace. Given branching event $e_b$,
%\begin{enumerate}
%	\item model the branching condition symbolically to execute along the negation of the original branch;
%	\item apply depth-first search (DFS) to uncover all execution suffixes under the unexplored branch (which may itself contain branching statements); and
%	\item for each suffix $t^s$,
%	\begin{enumerate}
%		\item truncate the original trace at $e_b$ yielding prefix $t_s$; and
%		\item concatenate $t_s$ with the negation of $e_b$ followed by $t^s$.
%	\end{enumerate}
%\end{enumerate}
%Constraint solving is applied to each of the resulting traces analogously to the original trace.

%We emphasize that exploration of new execution paths is subject to all the known limitations of symbolic execution, including in particular loop structures and object allocation. \tool\ currently fails if (i) the unexplored branch containts loops or (ii) there are object references that cannot be fully resolved at the branching point. In case of failure, \tool\ moves on to other branches. We demonstrate in Section \ref{sec:eval} that despite these limitations, the increase in coverage thanks to exploration of new paths is significant.

### 2014-09-03 overview.tex
tech

\begin{itemize}
\item $t$ refers to the thread that issues the event, denoted as $t^e$.
\item  $id$ refers to the id associated with each event, denoted as $i^e$. The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids. A natural id for each event is its index in the trace. 
\item $inst$ is the static instruction. The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.  Specifically, we are interested in the types of instructions listed in Table~\ref{Ta:syntax}. When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.  The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =$ in the branch. The target of the branch event is not important in our scope, therefore, we may omit the $goto$ part. The listed instructions are so basic that it can capture most events in a common imperative programming language. For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
\item $map$ maps the variable in the instruction to the runtime values. The map stores the value for each variable observed in the dynamic run. For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$. Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.
\end{itemize}

>>>>

\begin{itemize}
\item $t$ refers to the thread that issues the event, denoted as $t^e$.
\item  $id$ refers to the id associated with each event, denoted as $i^e$. The key property of  $id$ is {\em uniqueness}, i.e., any two events in the trace own different ids. The important property of $id$ is {\em monotonicity}, i.e.,  the events from the same thread should own the strictly increasing ids. Throughout this paper, we use  the index of an event in the trace as its id, which satisfies the above properties. 
\item $inst$ is the static instruction. The instructions are three-address instructions involving at most three operands, which modern compilers commonly support.  Specifically, we are interested in the types of instructions listed in Table~\ref{Ta:syntax}. When the variable does not appear on the left hand of an equation, such as $y$ in $x.f=y$, it may refer to a variable, a constant or event object creation expression $new (...)$.  The $bop$ stands for the binary operator, which may refer to $+, -, *, /, \%, \wedge, \vee$ in the assignment, or refer to $<, >, =, \wedge, \vee$ in the branch. The target of the branch event is not important in our scope, therefore, we may abbreviate the branch instruction as the boolean expression afterwards. The listed instructions are so basic that it can capture most events in a common imperative programming language. For example, the method call can be modeled by the above list, where the argument passing and value return are modeled as assignment events, and the virtual function resolution is modeled as the branch event conditioned on the type of the caller object.
\item $map$ maps the variable in the instruction to the runtime values. The map stores the value for each variable observed in the dynamic run. For simplicity, we use the notion $\lsyn x \rsyn^e$ to retrieve the value associated with the variable $x$ in $e$. 
\end{itemize}

%Specially, $\lsyn x.f \rsyn^e$ returns the location of $x.f$ (rather than  the value stored in the location), which is denoted as a pair $(\lsyn x \rsyn^e, f)$.

### 2014-09-03 trace.tex
tech

\tool\ handles the local assignment and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace and there is no need for the Phi node. More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event. The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once. Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition. 
\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread. Therefore we model them as local assignments: We introduce a fresh local variable {\tt l$_{\lsyn o.f \rsyn}^{id}$} to replace each definition and  the corresponding uses of a  field {\tt f} of the local object $o$. An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification (Section~\ref{sec:...}).
\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex. Each shared read may read from one of multiple writes that define the shared location, under different schedules.     To correlate the reads and the writes, we introduce two symbols, $W^{id}_{\tt \lsyn o.f\rsyn }$, which denotes write access to 
the field $f$ of the shared object $o$ by the event $id$, and $R^{id}_{\tt \lsyn o.f\rsyn }$,  which denotes the read access of the location $o.f$ by the event $id$. 
Similar to the local heap accesses, an assumption for the correlation is that the writes and reads still refer to the same base object $o$ in the predicted run, which we guarantee by specifying additional constraints (Section~\ref{sec:...}). 
\end{itemize}
\end{itemize}

@@

\begin{figure}
\centering
\begin{tabular}{l|l}
\hline
\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA form of  Trace$} \\
\hline
{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
\end{tabular}
\caption{Trace}
\label{fig:t4running2}
\end{figure}

@@

\subsection{Constraint System}

Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.

\paragraph{Intra-thread Order}

The first set of constraints reflects control flow within the individual threads, which must remain unchanged under the reordering transformation. Given input trace $t$, this is expressed as the following formula:
$$
\begin{array}{rl}
\forall \tau,\tau' \in t. & {\sf thread}\ \tau \equiv {\sf thread}\ \tau'. \\
& {\sf index}\ \tau < {\sf index}\ \tau' \Rightarrow O_{\tau} < O_{\tau'}
\end{array}
$$ 
The logical variables $O_x$ express ordering constraints. The requirement, as stated above, is that these variables
reflect the same order as the projection of ${\sf index}$ onto individual threads.

\paragraph{Race Condition}

Given pair $\tau$ and $\tau'$ of events that both access a common memory location $\ell$, we demand that
$$
\begin{array}{rl}
& {\sf thread}\ \tau \neq {\sf thread}\ \tau' \\
\bigwedge 	& (\ell \in {\sf writeset}\ \tau \vee \ell \in {\sf writeset}\ \tau') \\
\bigwedge   & O_{\tau} = O_{\tau'}
\end{array}
$$
That is, (i) events $\tau$ and $\tau'$ are executed by different threads, (ii) at least one of the events performs write access to $\ell$ (as judged by the ${\sf writeset}$ membership check), and (iii) the events occur simultanesouly. This is a direct logical encoding of the definition of a race condition.

\paragraph{Path Constraints}

The requirement with respect to branching is that the prefix of the original trace $t$ up to the pair $\tau$ and $\tau'$ of candidate racing events remains identical in the predicted trace $t'$. More accurately, under the assumption that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that 
$$
\bigwedge_{\tau'' \in t \cap {\bf bexp}.\
	\tau'' \in {\sf pre}\ \tau'} \lsyn {\sf stmt}\ \tau'' \rsyn\ t \equiv \lsyn {\sf stmt}\ \tau'' \rsyn\ t'
$$ 
where ${\sf stmt}$ is a helper function that obtains the code statement incident in a given transition. That is, all branching transitions up to $\tau'$ (which occurs after $\tau$) preserve their boolean interpretation under $t'$. This ensures that there are no divergences from the path containing the racing events, though beyond that path any feasible continuation is permitted. Importantly, contrary to \cite{JEFF-PLDI14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions is invariant under the input and predicted traces.

\paragraph{Variable Definitions}

A final requirement for the basic setting is that left-hand variables are defined according to the same right-hand variables as before. That is, version $i$ of variable ${\tt u}$ is defined as version $j$ of variable ${\tt v}$ in input trace $t$, then the same remains true in predicted  trace $t'$. This is enforced as the formula
$$
\bigwedge_{\tau'' \in t \cap {\bf asgn}.\
	\tau'' \in {\sf pre}\ \tau'} {\sf stmt}\ \tau'' \in t' 
$$
where we again assume that ${\sf index}\ \tau < {\sf index}\ \tau'$. That is, the same statement occurring in $t$ is also present in $t'$ (though the transitions may differ). Since the statements of $t'$ are a permutation of the statements of $t$, we are assured that use/def flow is constrained appropriately.

\paragraph{Relaxation of Flow Dependencies}

We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the data flow exhibited in the original trace. More precisely, \tool\ is able to relax flow dependencies in the original trace, whereby a thread reads a shared memory location written by another thread, while enforcing feasibility. This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same write access as in the original trace.

To ensure feasibility under relaxation of flow dependencies, we need to secure the link between the execution schedule and the write/read flow. As an illustration from the example in Figure \ref{fig:running}, $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule where thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}, the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}: $R_{\tt y}^4={\tt y}^1$.

The full and general constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
$$
\begin{array}{rll}
\bigvee_{e_w \in {\cal W}} &  & (R_{\ell} = {\ell}^{{\sf index}\ e_w}) \\
&		\bigwedge 	&  O(e_w) < O(e) \\
&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
\end{array}
$$
This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
($O(e) < O(e')$).

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at line {\tt 7} appears to match the write at line {\tt 5}, this is conditioned on the read at line {\tt 6} being linked to the assignment at line {\tt 4}. If the predicted run violates this link, then feasibility is no longer guaranteed. In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the write at line {\tt 5} no longer matches the read at line {\tt 7}.

\begin{figure}
	\centering
	\begin{tabular}{ll}
		\hline
		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
		\hline
		{\tt 1: x1 = new();} & \\
		{\tt 2: x2 = new();} & \\
		{\tt 3: y = x1;} & \\
		{\tt 4: y = x2;} & \\
		{\tt 5: x2.f = 5;} & \\	
		& {\tt 6: z = y;} \\
		& {\tt 7: w = z.f;} \\
	\end{tabular}
	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation}
\end{figure}

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced before the candidate racing events retain their original address in the predicted trace.
This achieves three guarantees: First, matching events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace. Finally, in a practical setting involving virtual method calls (which goes beyond our core grammar in Table \ref{Ta:syntax}), call-site resolutions are the same across the original and predicted traces. Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.

Formally, given pair $\tau$ and $\tau'$ of candidate racing events such that ${\sf index}\ \tau < {\sf index}\ \tau'$, 
we require that
$$
\bigwedge_{\tau'' \equiv {\tt y=x.f} \in t \cap {\bf heapr}.\
	\tau'' \in {\sf pre}\ \tau'} {\sf env}\ \sigma(t,\tau'')\ {\tt x} = {\sf env}\ \sigma(t',\tau'')\ {\tt x}
$$
where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
This constraint fixes that all heap dereferences up to the later of the candidate racing events retain their original base object as in the predicted trace $t'$.

>>>>

\tool\ handles the local assignment and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignment resembles the SSA form of static instructions in compiler optimization, except that the loops and recursions are fully resolved in a concrete trace and there is no need for the Phi node. More concretely, we replace the variable $v$ defined in an event, as well as the following uses of the definition, to a new variable $v^{id}$, where $id$ is the unique id of the event. The uniqueness of the id guarantees that no two events define the same variable, i.e., each variable is defined exactly once. Note that the uses of a definition can be computed easily by scanning the trace afterwards for the variables of the same name before the next redefinition. 
\item {\bf Heap accesses\ } The accesses of the local heap locations and the accesses of shared locations behave differently, therefore, we introduce different SSA form for them. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } The accesses of local heap locations behave similarly to the local assignment as both the definition and uses belong to the single thread. Therefore we model them as local assignments: We introduce a fresh local variable {\tt l$_{\lsyn o.f \rsyn}^{id}$} to replace each definition and  the corresponding uses of a  field {\tt f} of the local object $o$. An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification (Section~\ref{sec:...}).
\item  {\bf Shared Heap Accesses\ } The accesses of shared locations are more complex. Each shared read may read from one of multiple writes that define the shared location, under different schedules.     Given the write $x.f=y$ or the read $y=x.f$, we introduce two symbols, $W^{id}_{\tt x.f }$, which denotes the write access by the event $id$ to the field $f$ of the shared object referenced by $x$, and $R^{id}_{\tt  x.f }$,  which denotes the read access  by the event $id$ to the location $x.f$.  Note that, $x.f$ in the subscript is interpreted as the location, denoted as a pair $(\lsyn x \rsyn^e, f)$, rather than the value stored in the location. 
Similar to the local heap accesses, an assumption for the correlation is that the writes and reads still refer to the same base object $o$ in the predicted run, which we guarantee by specifying additional constraints (Section~\ref{sec:...}). 
\end{itemize}
\end{itemize}

@@

\begin{figure}
\centering
\begin{tabular}{l|l}
\hline
\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$SSA \ form \ of\   Trace$} \\
\hline
{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
\end{tabular}
\caption{Trace}
\label{fig:t4running2}
\end{figure}

@@

%TODO value or symbol for o.f, rethink about it.
\subsection{Constraint System}
Based on the SSA form of the trace, we build the constraints to compute a new trace with the new schedule and new mapping of variables to values.
To model the schedule, we introduce the order variable $O_{e}$ for each event $e$. Given two events $e_i, e_j$ from two different threads, $O_{e_i}<O_{e_j}$ in a trace means $e_i$ is scheduled before $e_j$. We omit the synchronization constraints purposely, which are well explained in all existing predictive analysis techniques~\cite{yannis, pldi14}.

%TODO synchronization constraint

%Having explained how the trace is encoded, we now describe in detail how constraints are derived from the trace, such that any permutation considered by the analysis is guaranteed to represent a feasible execution schedule.

{\bf Race Condition\ } Given any pair  of events, $e$ and $e'$,  that both access a common memory location $l$, we demand that
$$
\begin{array}{rl}
& t^e \neq t^{e'} \\
\bigwedge 	& (e\  writes\  l\   \vee e'\  writes\  l) \\
\bigwedge   & O_{e} = O_{e'}
\end{array}
$$
That is, (i) events $e$ and $e'$ are executed by different threads, (ii) at least one of the events performs write access to $l$, and (iii) the events run concurrently. This is a direct logical encoding of the definition of a race condition. We refer to both $e$ and $e'$ as the racy event.

{\bf Intra-thread Constraints\ } Given a racy pair under analysis, $e$ and $e'$, suppose $e$ is after $e'$ in the original trace $\tau$, the predictive analysis reschedules the events in the prefix of $e$, i.e., prior to $e$, so that $e$ and $e'$ can run concurrently. A necessary condition applied to predictive analysis techniques is that each thread should follow the same event sequence prior to $e$ as in the original run.  We explain the rationale underlying this requirement in Section~\ref{sec:proof}, but now proceed to explain how to guarantee it using constraints. The preservation of the event sequence for each thread requires the intra-thread constraints:

\begin{itemize}
\item {\bf Branch Decision Constraints\ } The branches in the prefix should take the same decisions to reproduce the same set of events. More formally,  
we require that 
$$
\bigwedge_{e_i \in \tau|\leq e \wedge  inst^{e_i}=x<y.\ }  x<y 	\equiv \lsyn x<y \rsyn^{e_i}
$$ 
where $e_i$ is a branch event of which the boolean expression is $x<y$ (we use the binary operator $<$ for illustration) and  $\lsyn x<y \rsyn^{e_i}$ stores the  evaluation result of the boolean expression in the original trace $\tau$. The constraint specifies that the boolean expression in the derived trace should be evaluated to the same boolean value. Importantly, different from existing analyses \cite{yannis,pldi14}, we do not pose the requirement that the values flowing into branching statements remain the same, but suffice with the relaxed requirement that the evaluation of branching expressions remain the same. For example, suppose  the branch event $x<y$ takes the value $1<2$ in the original trace, existing analyses require the same values for the variables $x$ and $y$ in the predicted run, while we allow  the branch event to take other values of the variables, such as $3<4$. 
\item {\bf Intra-thread Order Constraints\ } The events in the sequence should follow the same order as in the original trace. More formally, we require that
$$
\begin{array}{rl}
\forall e_i, e_j \in \tau. & t^{e_i}\equiv t^{e_j}. \\
& i^{e_i} < i^{e_j}\  \Rightarrow O_{e_i} < O_{e_j}
\end{array}
$$ 
The constraint specifies that the two events from the same thread should follow the same order as reflected by the ids of the events.
\item {\bf Assignment Constraints\ } The mapping of local variables should not contradict each instruction.   More formally, we require that
$$
\bigwedge_{e_i \in \tau|<=e \wedge type(e_i)=assign|heapr|heapw}\
	inst^{e_i}
$$
where $inst^{e_i}$, such as $x=y+z$ or $R^{id}_{x}=y$, captures the constraint over the values of the variables imposed by the instruction. For example, $x=y+z$ is not satisfied if the solver computes a mapping from the three variables to $3, 4, 5$ respectively. Again we consider only the events in the prefix $\tau|e$, where $e$ is the racy event that happens later. Specially, for the instruction involving object creation, $x=new (...)$, we encode the object as its address, i.e., a constant integer returned by $System.identityHashcode(x)$ at runtime.
\end{itemize}

{\bf Rescheduling through Relaxation of Flow Dependencies\ } We now move to the novel feature of \tool, which is its ability to explore execution schedules that depart from the value flow exhibited in the original trace. More precisely, \tool\ is able to relax value flow dependencies in the original trace, whereby a thread reads a
different value possibly written by other events (or even events from other threads) to the shared memory location, while enforcing feasibility.
 This is strictly beyond the coverage potential of existing predictive analyses, which restrict trace transformations to ones where any read access to a shared memory location must correspond to the same written values as in the original trace, which are usually from the same write events.

To ensure feasibility under relaxation of flow dependencies, we need to secure the flow between the read/write with the execution schedule. 
For example, in Figure \ref{fig:running}, for the read access to ${\tt y}$ at line {\tt 4} obtains the value assigned to ${\tt y}$ at line {\tt 1}, we need the schedule where  thread $T_1$ executes line {\tt 1}, then $T_2$ executes line {\tt 4}, and then the schedule switches again to $T_1$ to execute line {\tt 3}.

% $R_{\tt y}^4={\tt y}^1 \wedge O_1 < O_4 < O_3$ specifies that in a schedule , the $R_{\tt y}^4={\tt y}^1$.

%TODO l and ell
In general, the constraint formula, given read $R_{\ell}$ of location $\ell$ as part of event $e$ with set ${\cal W}$ of matching write events (i.e., events including write access to $\ell$), takes the following form:
$$
\begin{array}{rll}
\bigvee_{e_w \in {\cal W}} &  & (R^{e}_{\ell} = W^{e_w}_{\ell}) \\
&		\bigwedge 	&  O(e_w) < O(e) \\
&		\bigwedge_{e' \in {\cal W} \setminus \{ e_w \}} & (O(e') < O(e_w) \vee O(e) < O(e'))
\end{array}
$$
This disjunctive formula iterates over all matching write events, and demands for each that (i) it occurs prior to the read event ($O(e_w) < O(e)$) and (ii) all other write events either occur before ($O(e') < O(e_w)$) it or after the read event
($O(e) < O(e')$).

An important concern that arises due to relaxation of flow dependencies is that heap accesses may change their meaning, i.e., they involve different base objects and do no longer match with each other. As an illustration, we refer to Figure \ref{fig:heapAccess}. While the read at event {\tt 7} appears to match the write at event {\tt 5}, this is conditioned on the read at event {\tt 6} being linked to the assignment at event {\tt 4}. If the predicted run violates this link, then feasibility is no longer guaranteed. In particular, if reordering results in {\tt z} being assigned the first rather than second allocated object, then the read at event {\tt 7} no longer matches the write  at event {\tt 5}.

\begin{figure}
	\centering
	\begin{tabular}{ll}
		\hline
		\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
		\hline
		{\tt 1: x1 = new();} & \\
		{\tt 2: x2 = new();} & \\
		{\tt 3: y = x1;} & \\
		{\tt 4: y = x2;} & \\
		{\tt 5: x2.f = 5;} & \\	
		& {\tt 6: z = y;} \\
		& {\tt 7: w = z.f;} \\
	\end{tabular}
	\caption{\label{fig:heapAccess}Example illustrating the need to account for heap accesses during trace transformation, where the labelled number denotes the event id.}
\end{figure}

To address this challenge, we enhance the constraint system with the requirement that heap objects that are dereferenced in the following field reference before the candidate racing events retain their original address in the predicted trace.
This achieves two guarantees: First, matching heap access events in the original trace are guaranteed to also match in the predicted trace. Second, candidate races in the original trace remain viable in the predicted trace as they still refer to same location. Note that, the resolution of the virtual method call is modeled as the branch events that check the type of the target object, for which the constraints naturally guarantee the branch to take the same decision, i.e., the call to be resolved to the same method implementation.

%Finally, in a practical setting involving virtual method calls, call-site resolutions are the same across the original and predicted traces. 
%Notice that in this setting, we consider as relevant not only the targets of field dereferences but also the targets of virtual method invocations.

Formally, given pair $e$ and $e'$ of candidate racing events such that $e$ happens after $e'$ in the trace $\tau$, 
we require that
$$
\bigwedge_{ e_i \equiv {\tt y=x.f}|x.f=y \wedge e_i \in \tau|e }\
	x \equiv  \lsyn x \rsyn^{e_i}
$$
%where ${\sf env}$ is the state mapping from local variables to their value and $\sigma(t,\tau'')$ 
%(resp. $\sigma(t',\tau'')$) is the state arising at trace $t$ (resp. $t'$) immediately before event $\tau''$.
This constraint fixes that all heap dereferences  prior to the racy event retain their original base object as in the original trace $\tau$.

By sending the above constraints to a solver, we compute the necessary schedule orders among the events as well as the mapping of the variables. The necessary schedule orders define a partial order among the events, which permit a set of schedules that define the complete order that complies with it.
